---
title: "First Assignment"
author: "Juan David Larrota"
date: "Saturday, October 18, 2014"
output:
  html_document:
    keep_md: yes
---
## Loading and preprocessing the data
loading the data into R

```{r}
activity <- read.csv("~/Rstudio/Reproducible research/activity.csv")
no_na<- read.csv("~/Rstudio/Reproducible research/activity.csv")
```

## What is mean total number of steps taken per day?
Histogram of the total number of steps taken each day
```{r}
library(ggplot2)
totalSteps=tapply(activity$steps,activity$date,sum,na.rm=TRUE)
qplot(totalSteps, binwidth=1000, main= "total number of steps taken each day")
```

Mean and median total number of steps taken per day
```{r}
data.frame(mean(totalSteps),median(totalSteps))
```
## What is the average daily activity pattern?
time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
library(ggplot2)
aggdata=aggregate(activity$steps, by=list(activity$interval), mean, na.rm=TRUE)
colnames(aggdata)=c("interval", "steps")
qplot(interval,steps,data=aggdata,geom=c("line"), ylab="average number of steps taken",xlab=" 5-minute interval")
```

5-minute interval that contains the maximum number of steps?
```{r}
aggdata[which.max(aggdata$steps), ]
```

## Imputing missing values
total number of missing values in the dataset
```{r}
vec_NAS=is.na(activity$steps)
sum(vec_NAS)
vec_NAS=which(vec_NAS)
```
Fillimg NAS with the mean for the 5-minute interval
```{r}
for (i in 1:length(vec_NAS)){
 mean=as.numeric(tapply(no_na$steps, no_na$interval==no_na$interval[vec_NAS[i]], mean, na.rm=TRUE))
 no_na$steps[vec_NAS[i]]=mean[2]
}
```

histogram of the total number of steps taken each day and the mean and median total number of steps taken per day
```{r}
totalSteps=tapply(no_na$steps,no_na$date,sum,na.rm=TRUE)
qplot(totalSteps, binwidth=1000, main= "total number of steps taken each day")
data.frame(mean(totalSteps),median(totalSteps))
```
## Are there differences in activity patterns between weekdays and weekends?
```{r}
Sys.setlocale("LC_TIME", "English")
## Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
weekday.or.weekend <- function(date) {
  day <- weekdays(date)
  if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")) 
    return("weekday") else if (day %in% c("Saturday", "Sunday")) 
      return("weekend") else stop("invalid date")
}
no_na$date <- as.Date(no_na$date)
no_na$day <- sapply(no_na$date, FUN = weekday.or.weekend)
```
# Make a panel plot
```{r}
library(ggplot2)
aggdata1=with(no_na,aggregate(steps,by=list(day,interval),mean))
colnames(aggdata1)=c("day","interval","steps")
qplot(interval, steps,data=aggdata1, facets = .~day , geom="line")
```
